---
title: "R functions for data diagnostics"
author: "shelby bachman"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    toc: no
---
<hr>

```{r setup, include=FALSE}

rm(list=ls())
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = TRUE)

library(here)

source(here('check_collinearity.R'))
source(here('check_outliers.R'))
source(here('check_normality.R'))

```

Functions described below: 

- `check_normality()` / [source code](https://raw.githubusercontent.com/shelbybachman/useful-R-tools/master/check_normality.R)
- `check_outliers()` / [source code](https://raw.githubusercontent.com/shelbybachman/useful-R-tools/master/check_outliers.R)
- `check_collinearity()` / [source code](https://raw.githubusercontent.com/shelbybachman/useful-R-tools/master/check-collinearity.R)

These functions **perform checks of parametric assumptions** and **detect outliers**. These functions have a few dependencies which need to be installed but not loaded: `ggplot2` and `ggpubr`. To illustrate how to use these functions, I will use data from the `swiss` dataset that comes built-in with R:

```{r load_data, eval = TRUE, echo = FALSE}

head(swiss)

```

# check_normality()

This function **checks data contained in a vector for normality**. The function takes a vector as input and returns values of skewness (`output[[1]]`), kurtosis (`output[[2]]`), a message about whether severe non-normality is indicated (`output[[3]]`), and displays these values along with two plots (`output[[4]]`): a frequency distribution with the mean indicated with a dashed line, and a normal probability plot. It takes two mandatory arguments, a vector of values `data` and a name for the data `name`, as well as an optional argument `colour` specifying the color for the plots (it uses mint green by default).

```{r example_normality}

check_normality(data = swiss$Agriculture, name = 'Agriculture')[[4]]

```

# check_outliers()

This function **checks data contained in a vector for outliers**. The function detects outliers using the MAD-median rule, returns the row indices of those outliers (`output[[1]]`), and also displays a boxplot showing outliers identified according to the (non-robust) boxplot rule (`output[[2]]`). It takes two mandatory arguments, a vector of values `data` and a name for the data `name`, as well as an optional argument `colour` specifying the color for the plots (it uses mint green by default).

```{r example_outliers, eval = TRUE, echo = TRUE}

ggpubr::ggarrange(check_outliers(data = swiss$Agriculture, name = 'Agriculture')[[2]], 
                   check_outliers(data = swiss$Education, name = 'Education')[[2]])

```

# check_collinearity()

This function **checks two variables in a dataframe for extreme collinearity**. Given a dataframe and two variables of interest within that dataframe, it returns values of $R^2$ (`output[[1]]`), tolerance (`output[[2]]`), and a variance inflation factor (VIF) (`output[[3]]`), as well as these values superimposed on a scatterplot showing a linear regression line through the data (`output[[4]]`). It takes three arguments, a dataframe `data` as well as the names of two variables in the dataframe as the subsequent arguments.

```{r example_collinearity, eval = TRUE, echo = TRUE}

check_collinearity(swiss, 'Agriculture', 'Education')[[4]]

```



