---
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}

rm(list=ls())
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE)

library(rprojroot)

path <- function(x) find_root_file(x, criterion = has_file('useful-R-functions.Rproj'))
source(path('create-lm-table.R'))
source(path('create-lm-summary.R'))

```

# Create publication-ready tables summarizing a linear model fit

Use these functions if you have fit a linear model in R using the `lm` function and would like to compile the results into a concise set of metrics. To illustrate how to use these functions, I will use data from the `iris` dataset that comes built-in with R:

```{r load_data, eval = TRUE, echo = FALSE}

head(iris)

```

These functions have a few dependencies, which I load below:

```{r load_packages}

library(dplyr)
library(weights)
library(stringr)
library(tidyr)

```

Consider the case of a simple multiple linear regression analysis. For the `iris` dataset, I examine whether petal length and/or petal width are associated with sepal length:

```{r lm_overall}

lm.overall <- lm(Sepal.Length ~ Petal.Length + Petal.Width, data = iris)

```

The call to `create_lm_table` below generates a dataframe displaying the coefficients, their standard errors and confidence intervals, and associated *t*- and *p*-values. The first argument is the output of the call to the `lm` function. The second argument `include.intercept = 1` is included to indicate that I would like a row indicating these values for the intercept term. The third argument `vector.of.term.names` allows me to customize the names of the terms that will be displayed in the first column of the dataframe

```{r lm_overall_table}

create_lm_table(lm.object = lm.overall, 
                include.intercept = 1, 
                vector.of.term.names = c('Intercept', 'Petal Length', 'Petal Width'))

```

The call to `create_lm_summary` below generates a string summarizing the fit of the regression model. This function takes only one argument, the output of the call to the `lm` function: 

```{r lm_overall_summary}

create_lm_summary(lm.object = lm.overall)

```

Finally, I use the `kableExtra` package to combine these outputs into a single, publication-ready table:

```{r lm_overall_kable}

library(kableExtra)
table_overall <- create_lm_table(lm.object = lm.overall, 
                                 include.intercept = 1, 
                                 vector.of.term.names = c('Intercept', 'Petal Length', 'Petal Width'))
kable(table_overall) %>%
  kable_styling(full_width = F,
                'striped',
                position = 'left') %>%
  pack_rows(str_c(create_lm_summary(lm.object = lm.overall)),
            1, nrow(table_overall))

```

Next, I consider the case of a regression analysis with multiple groups, where I might want to create a table in which coefficients, etc. for each group are stacked on top of one another. I use the `iris` dataset again, this time performing a separate regression of petal length and width on sepal length, for each iris species separately:

```{r lm_groups}

lm.setosa <- lm(Sepal.Length ~ Petal.Length + Petal.Width, data = iris %>% filter(Species == 'setosa'))
lm.versicolor <- lm(Sepal.Length ~ Petal.Length + Petal.Width, data = iris %>% filter(Species == 'versicolor'))
lm.virginica <- lm(Sepal.Length ~ Petal.Length + Petal.Width, data = iris %>% filter(Species == 'virginica'))

```

I consolidate the steps above to create a single, composite table summarizing the output of all three model fits:

```{r lm_groups_kable}

table_setosa <- create_lm_table(lm.object = lm.setosa, 
                                 include.intercept = 1, 
                                 vector.of.term.names = c('Intercept', 'Petal Length', 'Petal Width'))
table_versicolor <- create_lm_table(lm.object = lm.versicolor, 
                                 include.intercept = 1, 
                                 vector.of.term.names = c('Intercept', 'Petal Length', 'Petal Width'))
table_virginica <- create_lm_table(lm.object = lm.virginica, 
                                 include.intercept = 1, 
                                 vector.of.term.names = c('Intercept', 'Petal Length', 'Petal Width'))

kable(as.data.frame(rbind(table_setosa,
                          table_versicolor,
                          table_virginica))) %>%
  kable_styling(full_width = F, 'striped', position = 'left') %>%
  pack_rows(str_c('Setosa, ', create_lm_summary(lm.object = lm.setosa), sep = ''), 1, nrow(table_setosa)) %>%
  pack_rows(str_c('Versicolor, ', create_lm_summary(lm.object = lm.versicolor), sep = ''), nrow(table_setosa)+1, nrow(table_setosa)+nrow(table_versicolor)) %>%
  pack_rows(str_c('Virginica, ', create_lm_summary(lm.object = lm.virginica), sep = ''), nrow(table_setosa)+nrow(table_versicolor)+1, nrow(table_setosa)+nrow(table_versicolor)+nrow(table_virginica))

```